---
title: "Multivariate Analysis"
description: |
  Study the different methods to visualise Multivariate Analysis
author:
  - name: Tan Yan Ru
    affiliation: MITB, Singapore Management University
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# 1. Install and Launch Packages Required

```{r}
packages = c('corrplot', 'tidyverse', 'ggstatsplot', 'ggcorrplot')

for(p in packages){library
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

# 2. Import and Prepare Data Set

```{r}
wine <- read_csv('data/wine_quality.csv')
```

# 3. Build Correlation Matrix: Using *pairs()* Method

## 3.1 Building Basic Correlation Matrix

Read all data from columns 1 to 11

```{r, echo=FALSE, eval=TRUE, fig.height=7, fig.width=7}
pairs(wine[,1:11])
```

As the top half and bottom half are the same, we will just need to show either the upper half or lower half of the matrix.

Below shows the method to show bottom half of the matrix.

```{r, echo=FALSE, eval=TRUE, fig.height=7, fig.width=7}
pairs(wine[,1:11], upper.panel = NULL)
```

Below shows the method to show upper half of the matrix.

```{r, echo=FALSE, eval=TRUE, fig.height=7, fig.width=7}
pairs(wine[,1:11], lower.panel = NULL)
```

Change the upper panel to other visualisation method.

```{r, echo=FALSE, eval=TRUE, fig.height=7, fig.width=7}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...) {
usr <- par("usr")
on.exit(par(usr))
par(usr = c(0, 1, 0, 1))
r <- abs(cor(x, y, use="complete.obs"))
txt <- format(c(r, 0.123456789), digits=digits)[1]
txt <- paste(prefix, txt, sep="")
if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
text(0.5, 0.5, txt, cex = cex.cor * (1 + r) / 2)
}

pairs(wine[,1:11], upper.panel = panel.cor)
```

## 3.2 Using ggstatsplot, ggcormat Package

### 3.2.1 Basic Plot

```{r, fig.height=7, fig.width=7}
ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11
)
```

### 3.2.1 Include Title, Sub-title

```{r, fig.height=7, fig.width=7}
ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11,
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  title    = "Correlogram for wine dataset",
  subtitle = "Four pairs are no significant at p < 0.05"
)
```

### 3.2.1 Multiple Plots

```{r, fig.height=15, fig.width=15}
grouped_ggcorrmat(
  data = wine,
  cor.vars = 1:11,
  grouping.var = type,
  type = "robust",
  p.adjust.method = "holm",
  plotgrid.args = list(ncol = 2),
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  annotation.args = list(
    tag_levels = "a",
    title = "Correlogram for wine dataset",
    subtitle = "The measures are: alcohol, sulphates, fixed acidity, citric acid, chlorides, residual sugar, density, free sulfur dioxide and volatile acidity",
    caption = "Dataset: UCI Machine Learning Repository"
  )
)
```

## 3.3 Using **corrplot** Package

### 3.3.1 Prepare Data for corrplot

Need to compute the correlation matrix of wine data frame

```{r}
wine.cor <- cor(wine[, 1:11])
```

Next, plot the corrplot()

```{r}
corrplot(wine.cor)
```

